## Abstract
The following analysis is performed on data from HAR [1]. The goal of the study is to construct model to predict type of exercise from gyroscopes' and accelerometers' readings.

I decide to use stochastic gradient boosting. It is more accurate for non-linear data than linear regression and faster than random forest and allows to fit model with multiple classes.

Data is too dense so I decide to use only 10% of it. Also I chose only non-unique variables with high variance as predictors.

The final model has accuracy about 90%.

## Analysis

### Setting up environment
```{r}
library(caret)
set.seed(114)

pml_write_files <- function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

```

### Training data loading and processing
```{r}
data <- read.csv("pml-training.csv")
# The dataset too big so I decided to reduce by sampling 10% of dataset
data <- data[sample(1:length(data[,1]), length(data[,1])/10),]

# The following vector of features has non-zero variance so I decided to use it for model fitting
data <- data[c("roll_belt", "pitch_belt", "yaw_belt", "gyros_belt_x", "gyros_belt_y", "gyros_belt_z", "accel_belt_x", "accel_belt_y", "accel_belt_z", "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm", "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x", "gyros_arm_y", "gyros_arm_z", "accel_arm_x", "accel_arm_y", "accel_arm_z", "magnet_arm_x", "magnet_arm_y", "magnet_arm_z", "total_accel_dumbbell", "gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z", "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", "magnet_dumbbell_x", "magnet_dumbbell_y", "magnet_dumbbell_z", "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm", "gyros_forearm_x", "gyros_forearm_y", "gyros_forearm_z", "accel_forearm_x", "accel_forearm_y", "accel_forearm_z", "magnet_forearm_x", "magnet_forearm_y", "magnet_forearm_z", "classe")]
data  <- data[complete.cases(data),]
classes <- unlist(data["classe"]); data["classe"] <- NULL
for (n in names(data)) {
  data[n] <- as.numeric(unlist(data[n]))
}
data["classe"] <- classes
```

### Creating validation set
```{r}
# I prefer 80% of data for training and 20% of data for validation
inTrain <- createDataPartition(y = data$classe, p = 0.8, list = FALSE)
training <- data[inTrain,]
testing <- data[-inTrain,]
```

### Model fitting
```{r, results='hide'}
plsFit <- train(training$classe ~ ., data = training, method = "gbm")
```

### Model diagnostics and validation
```{r}
confusionMatrix(predict(plsFit, newdata=testing), testing$classe)
```

### Testing data loading and processing
```{r}
pml_testing <- read.csv("pml-testing.csv")
pml_testing <- pml_testing[c("roll_belt", "pitch_belt", "yaw_belt", "gyros_belt_x", "gyros_belt_y", "gyros_belt_z", "accel_belt_x", "accel_belt_y", "accel_belt_z", "magnet_belt_x", "magnet_belt_y", "magnet_belt_z", "roll_arm", "pitch_arm", "yaw_arm", "total_accel_arm", "gyros_arm_x", "gyros_arm_y", "gyros_arm_z", "accel_arm_x", "accel_arm_y", "accel_arm_z", "magnet_arm_x", "magnet_arm_y", "magnet_arm_z", "total_accel_dumbbell", "gyros_dumbbell_x", "gyros_dumbbell_y", "gyros_dumbbell_z", "accel_dumbbell_x", "accel_dumbbell_y", "accel_dumbbell_z", "magnet_dumbbell_x", "magnet_dumbbell_y", "magnet_dumbbell_z", "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm", "gyros_forearm_x", "gyros_forearm_y", "gyros_forearm_z", "accel_forearm_x", "accel_forearm_y", "accel_forearm_z", "magnet_forearm_x", "magnet_forearm_y", "magnet_forearm_z")]
pml_testing  <- pml_testing[complete.cases(pml_testing),]
```

### Getting answers for submission
```{r}
answers <- as.character(predict(plsFit, newdata=pml_testing))
print(answers)
pml_write_files(answers)
```

## Citation
[1] Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.